---
title: "Reproducible Research - Peer Assignment 1"
output: html_document
---

First we load in the data and convert to date:

```{r} 
  library(dplyr)
  'Load in data'
  activity <- read.csv(file = 'activity.csv', header = TRUE)
  activity$date <- as.Date(activity$date, format = '%Y-%m-%d')
```  
  
Question 1: What is the mean # of total steps taken per day

```{r}  
  total_steps <- summarize(group_by(activity, date), s = sum(steps, na.rm = TRUE))
  
  'Plot the histogram'
  hist(total_steps$s)
  
  'Mean & Median of data'
  steps_mean <- summarize(total_steps, mean(s))
  steps_median <- summarize(total_steps, median(s))
```

Question 2: What is the average daily activity pattern?

```{r}
  'Processing of data'
  steps_pattern <- summarize(group_by(activity, interval), 
                             mean_Steps = mean(steps, na.rm = TRUE))
  
  'plot a line graph of the average day'
  plot(steps_pattern, type = 'l')
  
  'find the interval where the avg steps is the greatest'
  steps_pattern$interval[steps_pattern$mean_Steps == max(steps_pattern$mean_Steps, na.rm = TRUE)]
```
  

Question 3: Imputing missing values

1. How many rows have NA:
```{r}
  sum(is.na(activity$steps))
```

2. Strategy: Use the mean of the interval for each NA value

3. Input the mean of the interval for each NA
```{r}
  j <- length(activity$steps)
  new_Activity <- activity
  for (i in which(is.na(activity$steps)))
  {
    new_Activity$steps[i] <- steps_pattern$mean_Steps[i]
  }
```

4. Create a new histogram
```{r}
  new_steps <- summarize(group_by(new_Activity, date), s = sum(steps, na.rm = TRUE))
  hist(new_steps$s)
```


Question 4: Are there differences in activity patterns between weekdays and weekends?

1. Create new factor with weekdays / weekends
```{r}
  
  day <- weekdays(new_Activity$date)
  week <- mutate(new_Activity, week = "")
  
  for (i in 1:length(day))
  {
    if(day[i] == 'Saturday' | day[i] == 'Sunday')
      week$week[i] <- 'Weekend'
    else
      week$week[i] <- 'Weekday'
  }
```
  
2. Create the plots to determine differences  
```{r}
  
  new_steps <- summarize(group_by(week, interval, week), s = sum(steps, na.rm = TRUE))
  plot(new_steps$s[new_steps$week == 'Weekday'], type = 'l')
  plot(new_steps$s[new_steps$week == 'Weekend'], type = 'l')
```
  
